Trends of SDR in Thessaloniki

### 2023-07-03 Article.Rmd
Commit 2023-07-03 12:17

## Long-term trends

Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
Only for All-sky (Fig. \ref{fig:trendALL}) <!--and Cloudy-sky (Fig. \ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab. \ref{tab:trendtable}), to draw some conclusions.

```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
knitr::include_graphics("./images/LongtermTrends-2.png")
```

@@

<!-- \footnotesize -->
```{r trendtable, echo=F}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")

panderOptions("table.alignment.default", "center")
pander(LT,
       caption = "\\label{tab:trendtable}Trends of daily means by sky conditions.")
# \\label{tab:tab2}
# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")

## some vars to use
alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
```
<!-- \normalsize -->

>>>>

## Long-term SDR trends

We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table \ref{tab:trendtable}).

```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`. The black line shows the long term trend for all-sky conditions."}
knitr::include_graphics("./images/LongtermTrends-2.png")
```

@@

\footnotesize
```{r trendtable, echo=F}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
## no need all are 100%
LT$`Trend statistical signif. [%]` <- NULL

panderOptions("table.alignment.default", "center")
pander(LT,
       caption = "\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above $99.9\\%$")
# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")

## some vars to use later
alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
```
\normalsize

### 2023-07-03 Article.tex
Commit 2023-07-03 12:17

  \title{Trends of SDR in Thessaloniki}
    \author[Laboratory of Atmospheric Physics]{Natsis Athanasios%
  \corref{cor1}%
  }
   \ead{natsisphysicist@gmail.com} 
    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
  %
  }
   \ead{abais@auth.gr} 
      \affiliation[Laboratory of Atmospheric Physics]{
    organization={Aristotle University of
Thessaloniki},addressline={Campus Box
149},city={Thessaloniki},postcode={54124},country={Greece},}
    \cortext[cor1]{Corresponding author}

@@

\hypertarget{data-and-methodology.}{%
\section{Data and methodology.}\label{data-and-methodology.}}

Location

The SDR data used in this study cover the period from January 1993 to
March 2011, and were measured with a Kipp \& Zonen CM-21 pyranometer
operating contin- uously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\$40\^{}\circ\$38''N,
\$22\^{}\circ\$57''E, ∼80 m above sea level). The monitoring site is
located near the city centre. The horizon is clear of obstruc- ◦ ◦ tions
except for the azimuth angles between 35 and 120where obstacles of up to
∼10◦ elevation block the direct irradiance in the morning for SZA larger
than ∼80◦ . These data points were excluded from the analysis, as

Instruments and callibrations

relation to aerosols

Beijing (40 N, 116 E), which are representative of regions with high
aerosol load. Northern Greece has been selected because it is well known
to be located at the crossroads of aerosols and various pollutants
pathways originating in central and Eastern Europe (Lelieveld et al.,
2002; Zerefos et al., 2002). Thessa- loniki, the major city in northern
Greece, lies along the exit of air pollutants from Europe, in the path
of the prevailing long-range transport (LRT) trajectories, as we move
from central and east- ern Europe to the Mediterranean (Zerefos et. al.,
2002). On the

The SDR is equivalent to the whole sky Global radiation, also refered as
Global Horizontal Irradiance (GHI), and was obtained with a horizontal
leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
2023-03-31. In some algorithms we used as auxiliary data, the direct
beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
with data availability from 2016-04-01 to 2023-03-31.

\ldots{}

\ldots{}

There are three distinct steps to the creation of this dataset: a) the
acquisition of radiation measurements from the sensors, b) a radiation
data quality check, and c) the identification of ``clear sky''
conditions from the radiometric data d) data aggregation and trend
analysis.

For the acquisition of radiometric data, the signal of the broadband
instruments is sampled with a rate of \(1 \text{Hz}\). The mean and
standard deviation values are recorded for every minute. The
measurements are corrected for the zero offset of the instrument signal.
As reference of the ``dark signal'' is used the Sun elevation angle
bellow \(-10^\circ\), for a period of \(3 \text{h}\). The signal is
converted to radiation flux, using a ramped value of the instrument
sensitivity, derived from the eight laboratory calibrations of the
instrument, during the study period.

A manual screening is performed, to remove inconsistencies and erroneous
recordings, that can occur randomly or systematically, during the long
continuous operation of the instrument. The manual screening is aided by
a radiation data quality assurance procedure, adjusted for this site,
based on methods of Long and Shi~\citetext{\citealp[
]{Long2008a}; \citealp{Long2006}}. Thus, problematic recordings have
been excluded from further processing. Although it is not possible to
detect all the bad data, the large number of data and the aggregation
scheme we use, can provide us with accurate radiometric measurements,
for the scope of this study.

\hypertarget{clear-sky-identification}{%
\subsection{Clear sky identification}\label{clear-sky-identification}}

@@

Our method to characterize the sky conditions and the definition of each
sky condition has some subjectivity. The algorithm was calibrated with
the main focus to identify the presence of clouds on the sky dome,
although there are marginal cases that there will be false positives or
false negatives identifications.

@@

\hypertarget{mean-value-of-irradiance-during-the-time-period.}{%
\subsubsection{Mean value of irradiance during the time
period.}\label{mean-value-of-irradiance-during-the-time-period.}}

@@

\hypertarget{max-value-of-irradiance-during-the-time-period.}{%
\subsubsection{Max value of irradiance during the time
period.}\label{max-value-of-irradiance-during-the-time-period.}}

@@

\hypertarget{variability-in-irradiance-by-line-length.}{%
\subsubsection{Variability in irradiance by line
length.}\label{variability-in-irradiance-by-line-length.}}

@@

\hypertarget{variance-of-changes-in-the-time-series.}{%
\subsubsection{Variance of Changes in the Time
series.}\label{variance-of-changes-in-the-time-series.}}

@@

\begin{gather}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
\end{gather}

@@

\hypertarget{variability-in-the-shape-of-the-irradiance-measurements.}{%
\subsubsection{Variability in the Shape of the irradiance
Measurements.}\label{variability-in-the-shape-of-the-irradiance-measurements.}}

@@

In order to estimate the effect of the clouds on the SDR we created
three datasets, by characterizing each one-minute measurement with a
corresponding sky condition. The all-sky conditions, containing all the
valid measurements. The clear-sky conditions data, where we have
inferred that the sky was almost clear of clouds, and the remainder part
the data as cloudy sky conditions data (cloud-sky). To identify the
clear-sky conditions we used the method proposed by \citet{Long2000} and
by \citet{Reno2016}, that was adapted and configured for the site. We
have to note, that the definition of what constitutes as clear or cloudy
sky, has some subjectivity, due to the used method of characterization.
As a result, the details of the definition are site specific, it relies
on a combination of threshold of comparisons with ideal actinometric
models and statistics on different signal behavior.

\hypertarget{data-and-data-selection.}{%
\subsection{Data and data selection.}\label{data-and-data-selection.}}

@@

\hypertarget{long-term-trends}{%
\subsection{Long-term trends}\label{long-term-trends}}

Using the mean daily SDR we produce the trends of each data set, as
departure from the seasonal value. Only for All-sky (Fig.
\ref{fig:trendALL}) conditions, the statistical significance of the
trends are acceptable (Tab. \ref{tab:trendtable}), to draw some
conclusions.

@@

{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-2}

@@

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.3996 & 8.665e-23 & All sky cond. & 100 \\
0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
\end{longtable}

@@

{\centering \subfloat[Trend distribution for all-sky conditions.(\#fig:szatrends-1)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.(\#fig:szatrends-2)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }

@@

\caption{Distribution of the SDR's long term trends by SZA.}(\#fig:szatrends)
\end{figure}

@@

\caption{Trends by season for all condition. Displaying monthly means of daily means.}(\#fig:seasonalALL)
\end{figure}

@@

{\centering \subfloat[one plot(\#fig:cumsum-1)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one(\#fig:cumsum-2)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }

@@

\caption{two plots}(\#fig:cumsum)
\end{figure}

>>>>

  \title{Trends of SDR in Thessaloniki}
    \author[Laboratory of Atmospheric Physics]{Athanasios Natsis%
  \corref{cor1}%
  }
   \ead{natsisphysicist@gmail.com} 
    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
  %
  }
   \ead{abais@auth.gr} 
      \affiliation[Laboratory of Atmospheric Physics]{
    organization={Aristotle University of
Thessaloniki},addressline={Campus Box
149},city={Thessaloniki},postcode={54124},country={Greece},}
    \cortext[cor1]{Corresponding author}

@@

\hypertarget{observational-data-and-methodology.}{%
\section{Observational data and
methodology.}\label{observational-data-and-methodology.}}

The SDR is equivalent to the whole sky Global radiation (or Global
Horizontal Irradiance, GHI), and was obtained with a horizontal leveled,
CM-21 pyranometer, from 1993-04-12 to 2023-03-31. Also, we have
measurements of direct beam radiation (DNI) obtained by a collocated
CHP-1 pyrheliometer, from 2016-04-01 to 2023-03-31. Although, we
performed a similar analysis to the DNI as for the SDR, the results are
not presented here, as they lack the appropriate statistical
significance, due to the sorter duration of the data. Moreover, the DNI
was used as auxiliary data, with the clear sky identification algorithm
(CSid), to aid with the selection of the appropriate thresholds. We have
to note also, that despite the capability of the CSid to use the DNI as
an characterization parameter, we haven't use it, to avoid any selection
bias due to unequal availability of measurements.

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ38'\)N,
\(22^\circ57'\)E, 80m a.s.l.). The monitoring site is located near the
city centre. During the study period the CM-21 pyranometer has been
independently calibrated four times. There are three distinct steps to
the creation of this dataset: a) the acquisition of radiation
measurements from the sensors, b) a radiation data quality check, and c)
the identification of ``clear sky'' conditions from the radiometric data
d) data aggregation and trend analysis.

For the acquisition of radiometric data, the signal of the broadband
instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its
standard deviation are recorded for every minute. The measurements are
corrected for the zero offset of the instrument signal. As reference of
the ``dark signal'' is used the Sun elevation angle bellow
\(-10^\circ\), for a period of \(3 \text{h}\). The signal is converted
to radiation flux, using a ramped value of the instrument sensitivity,
derived from the eight laboratory calibrations of the instrument, during
the study period.

A manual screening was performed, to remove inconsistencies and
erroneous recordings, that can occur randomly or systematically, during
the long continuous operation of the instrument. The manual screening is
aided by a radiation data quality assurance procedure, adjusted for this
site, based on methods of Long and
Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
problematic recordings have been excluded from further processing.
Although it is not possible to detect all the bad data, the large number
of data and the aggregation scheme we use, can provide us with accurate
radiometric measurements, for the scope of this study.

In order to estimate the effect of the clouds on the SDR, we created
three datasets, by characterizing each one-minute measurement with a
corresponding sky condition. The all-sky conditions, containing all the
valid measurements. The clear-sky conditions data, correspond to data
where we have inferred that the sky was almost clear of clouds, and the
remainder part the data was characterized with cloud sky conditions
(cloud-sky). To identify the clear-sky conditions we used the method
proposed by \citet{Long2000} and by \citet{Reno2016}, that was adapted
and configured for the site.

cite iterative process

@@

We have to note, that the definition of what constitutes as clear or
cloudy sky, has some subjectivity, due to the used method of
characterization. As a result, the details of the definition are site
specific, it relies on a combination of threshold of comparisons with
ideal actinometric models and statistics on different signal behavior.

The algorithm was calibrated with the main focus to identify the
presence of clouds on the sky dome, although there are marginal cases
that there will be false positives or false negatives identifications.

@@

@@

@@

@@

@@

\begin{equation}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
\end{equation}

@@

@@

@@

\hypertarget{long-term-sdr-trends}{%
\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}

We calculated the SDR trends, from the mean daily values of the
departure from the seasonal values, for the whole time series for the
three sky conditions (Table \ref{tab:trendtable}).

@@

{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}

@@

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
\end{figure}

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
\caption{\label{tab:trendtable}Trends of daily SDR means, for different
sky conditions. All cases have a statistical significance above
\(99.9\%\) \normalsize}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.3996 & 8.665e-23 & All sky cond. \\
0.3773 & 7.075e-18 & Clear sky cond. \\
0.3594 & 3.318e-09 & Cloudy cond. \\
\end{longtable}

@@

{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }

@@

\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
\end{figure}

@@

\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
\end{figure}

@@

{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }

@@

\caption{two plots}\label{fig:cumsum}
\end{figure}
